The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2017-08-09 JStatSoft_forestinventory_master.rnw
update Andreas 03.08.2017(2)
update Andreas 09.08.2017
update Andreas 09.08.2017(4)

@@

\SweaveInput{rnw/Global Estimators and their Application(2).Rnw}

\SweaveInput{rnw/SAE Estimators and their Application.Rnw}

@@

@@

>>>>

% execute everything that goes without echo:
<<echo=FALSE>>=
library(forestinventory)
@

@@

\SweaveInput{rnw/Twophase_Estimators and their Application.Rnw}

\SweaveInput{rnw/Threephase_Estimators and their Application.Rnw}

@@

\SweaveInput{rnw/Calculation of Confidence Intervals.Rnw}

@@

\SweaveInput{rnw/Special Cases and Scenarios.Rnw}

### 2017-08-09 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 09.08.2017(4)

\section{Twophase Estimators and their Application}
\label{sec:twophase_and_appl}

% ---------------------------------------------------------------------------- %
\subsection{Global Estimators}

% NOTES:
% - The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (so called \textit{onephase estimation})
%   by the \code{onephase} function in \pkg{forestinventory}. This function can also be used to access the gain in efficiency of the mutliphase estimation procedures. 
%
% - we only use the non-cluster sampling here to illustrate:
% a) reg-coef calculation, point estimates and two variances, important components
% b) we introduce the boundary adjustments: explain why, give mathematical formula, mention that this is optional and can be left out, in which case the simple means of Z are used
% c) exhaustive and non-exhaustive

% Here, we give the \textbf{mathematical background} of:
% \begin{itemize}
%   \item the classical two-phase estimator, including:
%     \item the external and g-weight variance. Explain the differences and pros for using the g-weight version
%     \item the auxiliary components (Z) and covvar(Z) as well as their use in the point and variance estimator
%   \item the boundary weight adjustment (mathematically: weighted means of Z) --> our extension to the already published stuff
%   \item ...
% \end{itemize}
% 
%

% 
% Here, we give the \textbf{application} examples:
% \begin{itemize}
%   \item example for non-exhaustive case: use grison dataset
%     \item with boundary adjustments, just mention that this is optional and can be left out, in which case the simple means of Z are used
%   \item example for exhaustive case: use cluster set
% \end{itemize}

\subsubsection{Mathematical Background}

The vector regression coefficients of the OLS regression model are found by solving the sample-based normal equation.

\begin{equation}\label{normequ_simple}
  \hat{\pmb{\beta}}_{s_2}&=& \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1} \Big(\frac{1}{n_2}\sum_{x\in{s_2}}Y(x)\pmb{Z}(x)\Big)
\end{equation}

This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables. Based on the regression coefficients, the prediction for the sample location $x$ is calcualted as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$. Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations. This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase} and \code{threephase} function as input data. \citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g. from previous studies. However, a respective function in the package is up to now non existing and will be subject to future extentions.\par

The design-based variance-covariance matrix of the regression coefficients $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$ is then calculated as

\begin{equation}\label{eq:estvarmatrix}
  \hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}:=\Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
  \Big(\frac{1}{n_2^2}\sum_{x\in{s_2}}\hat{R}^2(x)\pmb{Z}(x)\pmb{Z}(x)^t\Big)
  \Big(\frac{1}{n_2}\sum_{x\in{s_2}}\pmb{Z}(x)\pmb{Z}^t(x) \Big)^{-1}
\end{equation}

The \textbf{point estimates} and their \textbf{variances} can now be calculated for the case of \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables. Note that exhaustively available auxiliary information does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information. Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios. In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$. The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in figure \ref{fig:exh_nexh_bw}. In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}). This allows for calculating the \textit{true} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$ and for using this information in the estimators. In the \textit{non-exhaustive} estimator case, the explanatory variables are not computed at every sample location, but only at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}). Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (see equation \ref{eq:gw_var_2p_reg_nex}). As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption must be relaxed.\par

An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{boundary adjustment} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package. In forest inventories, the sample is often restricted to those sample locations located within the forest area. In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g. by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}). This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy. To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}. The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.

\begin{equation}\label{eq:wmeanZ}
  \hat{\bar{\pmb{Z}}}=\frac{\sum_{x\in{s_1}}w(x)\bmp{Z}(x)}{\sum_{x\in{s_1}}w(x)}
\end{equation}

The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively. Note that this form results under particular case where the regression coefficients are derived using the data from the current inventory (i.e. \textbf{internal} model). In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}R(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.

\begin{subequations}\label{eq:pointest_2p_reg}
\begin{align}
  \hat{Y}_{reg2p,ex} & =\bar{\pmb{Z}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_ex}\\
  \hat{Y}_{reg2p,nex} & =\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_nex}
\end{align}
\end{subequations}

The package calculates two kind of variances for the point estimate. The estimated \textbf{design-based variance} is given in equation \ref{eq:gw_var_2p_reg} for the exhaustive (equation \ref{eq:gw_var_2p_reg_ex}) and non-exhaustive extimator (equation \ref{eq:gw_var_2p_reg_nex}). Note that this is mathematically identical to the \textbf{g-weight} formulation of the design-based variances given in \citep{mandallaz2016}.

\begin{subequations}\label{eq:gw_var_2p_reg}
\begin{align}
  \hat{\var}(\hat{Y}_{reg2p,ex}) & :=\hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}} \label{eq:gw_var_2p_reg_ex}\\
  \hat{\var}(\hat{Y}_{reg2p,nex}) & :=
  \hat{\bar{\pmb{Z}}}^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}
  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}}\hat{\pmb{\beta}}_{s_2} \label{eq:gw_var_2p_reg_nex}
\end{align}
\end{subequations}

where the uncertainty caused by the estimation of $\bar{\pmb{Z}}$ by $\hat{\bar{\pmb{Z}}}$ is considered by the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}$, estimated by
\begin{equation}\label{estvarcovaux}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{1}}=
\frac{1}{n_{1}(n_{1}-1)}\sum_{x\in{s_{1}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{1})^t
\end{equation}

The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}). Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.

\begin{subequations}\label{eq:varexternal_2p_reg}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(Y(x)-\bar{Y}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & =
 \frac{1}{n_1}\frac{1}{n_2-1}\sum_{x\in{s_2}}(Y(x)-\bar{Y}_2)^2+
(1-\frac{n_2}{n_1})\frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(R(x)-\bar{R})^2 \label{eq:varexternal_2p_reg_nex}
\end{align}
\end{subequations}

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{1.12\hsize}{!}{\includegraphics{fig/exh_and_boundweights(2).PNG}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
   \resizebox{0.77\hsize}{!}{\includegraphics{fig/nexh_and_boundweights.PNG}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_b}
	\end{subfigure}
\caption{Concept of using exhaustive (\textbf{(a)}) and non-exhaustive (\textbf{(b)}) auxiliary information including boundary adjustment. Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle. A set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within the small squares at each sample location $x$ (points) that falls into the forest area (underlying polygon). w: weights used for calculating the weighted mean of each explanatory variable. Dotted circles indicate terrestrial sample plots.}
\label{fig:exh_nexh_bw}
\end{figure}

\subsubsection{Application}

To demonstrate the use of the global two-phase estimators, we will use the \code{grisons} data frame that comes with installing the package from the CRAN repository. The dataset contains data from a two-phase forest inventory under simple sampling design conducted in 2007 that was used in \citet{mandallaz2013b} and comprises observations from 306 sample locations ($s_1$ sample). Auxiliary information for all 306 plots is provided in the form of LiDAR canopy height metrics (\code{mean}, \code{stddev}, \code{max}, \code{q75}). For a systematic subsample of 67 out of the 306 plots, terrestrial information of the timber volume per hectare (\code{tvol}) on the sample plot level is provided from a terrestrial survey ($s_2$ sample). We can load the data frame in the \proglang{R} environment by executing:

\begin{small}
<<>>=
data("grisons", package = "forestinventory")
@
\end{small}

Note that the structure of the data frame has to meet certain requirements to be used by the \code{twophase} and \code{threephase} functions of the package. Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location. Note that for global two-phase estimation, the data frame has to comprise a column that indicates whether a row corresponds to a $s_1$ or $s_2$ sample location (in \code{grisons}, this is the column \code{phase_id_2p}).\par

In order to apply the \textbf{non-exhaustive two-phase estimation with boundary weight adjustment} to the \code{grisons} data, we have to specify a) the regression model (\code{formula}) as specified in the \code{lm}-function \citep{R}, b) the data frame containing the inventory information (\code{data}), c) the name of the column specifying to which sample ($s_1$, $s_2$) an observation belongs (\code{phase.col}) as well as the indicator for the $s_2$, i.e. terrestrial sample (\code{terrgrid.id}), and d) the columname of the column containing the weights $w(x)$ of the boundary adjustments. The \textbf{non-exhaustive estimator with boundary weight adjustment} can thus be applied as follows:
\begin{small}
<<>>=
reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
                   data=grisons,
                   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
                   boundary_weights = "boundary_weights")
@
\end{small}
The \code{twophase}-function creates an object of class \code{twophase}. A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object. It first provides a summary of the function call, the applied regression formula and the estimation method that was applied. The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}). Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model.
\begin{small}
<<>>=
summary(reg2p_nex)
@
\end{small}
If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}. The vector must be input in the same order that the \code{lm}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1. We assume that in this case, the boundary adjustment has been applied when calculating the exact mean of $\pmb{Z}$.

The right order can easily be established by extracting the columnames of the respective \code{lm} output. This step is particularly recommended if categorical variables are used. In our case this is:
\begin{small}
<<eval=FALSE>>=
colnames(lm(formula = tvol ~ mean + stddev + max + q75, data = grisons, x = TRUE)$x)
@
\end{small}
The \textbf{exhaustive estimator} can thus be applied after creating the vector of exact means $\bar{\pmb{Z}}$:
\begin{small}
<<>>=
true.means.Z <- c(1, 11.39, 8.84, 32.68, 18.03)

reg2p_ex <- twophase(formula = tvol ~ mean + stddev + max + q75,
                    data = grisons,
                    phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                    exhaustive = true.means.Z)
@
\end{small}

We can have a look on the estimation results by particularly querying the \code{estimation} object of \code{reg2p_ex}:
\begin{small}
<<>>=
reg2p_ex$estimation
@
\end{small}
Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation. This is because we took out one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.

%--------------------------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------------------------%

\subsection{Small Area Estimators}

\subsubsection{Mathematical Background}

The package \pkg{forestinventory} offers 3 small area estimators. A general principle of all small area estimators of the package is to restrict the information to the small area $G$ whenever possible. An exception however is the retrieval of the regression coefficient vector $\hat{\pmb{\beta}}$ and its variance-covariance matrix $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$, which is always calculated using the entire sample $s_2$ according to equation \ref{normequ_simple} and \ref{eq:estvarmatrix}.\par

%----------------------------------------------- %
% \textbf{Small and Pseudo Small Area Estimator}\par

We first introduce the \textbf{small area estimator} (\textit{small}), which uses exhaustively computed explanatory variables, and its non-exhaustive version, the \textbf{\textit{pseudo}} \textbf{small area estimator} (\textit{psmall}). Their point estimates are given in equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}. As obvious, the globally derived regression coefficients are first applied to the exhaustively or non-exhaustively calculated means of the explanatory variables available at the sample location $s_{1,G}$ in the small area. A potential bias of the regression model predictions in the small area $G$ is then corrected by adding the mean of the residuals in $G$ to the first term. We call this the \textit{bias} or \textit{residual correction} term.

\begin{subequations}\label{eq:pest_2p_small_psmall}
\begin{align}
  \hat{Y}_{G,small,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x)  \label{eq:pointest_2p_reg_small} \\
  \hat{Y}_{G,psmall,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} + \frac{1}{n_{2,G}}\hat{R}(x) \label{eq:pointest_2p_reg_psmall}
\end{align}
\end{subequations}

The package again provides the \textit{design-based} (equation \ref{eq:var_2p_reg_small}, \ref{eq:var_2p_reg_psmall}) and \textit{external} variance (equation \ref{eq:varext_2p_reg_small}, \ref{eq:varext_2p_reg_psmall}) for the \textbf{small area estimator} and the \textbf{pseudo small area estimator}. Again note that all components are restricted those available at the sample locations $s_{1,G}$ and $s_{2,G}$ in the small area, with exception of the regression coeffient components $\hat{\pmb{\beta}}_{s_2}$ and $\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}$.

\begin{subequations}\label{eq:var_2p_small_psmall}
\begin{align}
  \hat{\var}(\hat{Y}_{G,small,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G 
    + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_small} \\
  \hat{\var}(\hat{Y}_{G,psmall,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2} 
  + \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:var_2p_reg_psmall}
\end{align}
\end{subequations}

\begin{subequations}\label{eq:varext_2p_small_psmall}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2  \label{eq:varext_2p_reg_small} \\
  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(Y(x)-\bar{Y}_{2,G}\Big)^2 \nonumber \\ 
 &+ \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\hat{\bar{R}}_{2,G}\Big)^2 \label{eq:varext_2p_reg_psmall}
\end{align}
\end{subequations}

The the variance-covariance matrix of the auxiliary vector $\hat{\bar{\pmb{Z}}}_G$ is estimated by
\begin{equation}\label{estvarcovaux_G}
\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_{G}}=
\frac{1}{n_{G}(n_{G}-1)}\sum_{x\in{s_{1,G}}}
(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})(\pmb{Z}(x)-\hat{\bar{\pmb{Z}}}_{G})^t
\end{equation}

%------------------------------------------------------------- %
% \textbf{Synthetic and Pseudo Synthetic Estimator}\par

The \textbf{synthetic} and \textbf{pseudo synthetic estimator} is applied if no terrestrial samples are available within the small area $G$ (i.e. $n_{2,G}=0$). In this case, the point estimate (equation \ref{eq:pointest_2p_reg_synth} and \ref{eq:pointest_2p_reg_psynth}) is based \textit{only} on the predictions generated by applying the globally derived regression model to the auxiliary vector $\bar{\pmb{Z}}_G$ or $\hat{\bar{\pmb{Z}}}_G$. However, the bias correction applied in the small and pseudo small area estimator (equation \ref{eq:pointest_2p_reg_small} and \ref{eq:pointest_2p_reg_psmall}) is no longer possible. Thus, the synthetic / pseudo synthetic estimator is unbiased \textbf{only} if the \textit{zero mean property} of the residuals $\int_F R(x)dx=0$ also holds for the residuals in the small area $G$, i.e. $\int_G R(x)dx=0$. If this assumption is not fullfilled, the synthetic point estimate $\hat{Y}_{G,synth,2p}$ and the pseudo synthetic point estimate $\hat{Y}_{G,psynth,2p}$ will have a design-based asymptotic bias equal to $-\frac{1}{\lambda(G)}\int_G R(x)$. Also note that residual variation can no longer be considered in the design-based variance (equation \ref{eq:var_2p_reg_synth} and \ref{eq:var_2p_reg_psynth}). Therefore, the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias. Due to the absence of available residuals in $G$, there is also no external variance form for synthetic estimations.

\begin{subequations}\label{eq:pest_2p_synth_psynth}
\begin{align}
  \hat{Y}_{G,synth,2p} & =\bar{\pmb{Z}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_synth} \\
  \hat{Y}_{G,psynth,2p} & =\hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\beta}}_{s_2} \label{eq:pointest_2p_reg_psynth}
\end{align}
\end{subequations}

\begin{subequations}\label{eq:var_2p_synth_psynth}
\begin{align}
  \hat{\var}(\hat{Y}_{G,synth,2p}) & := \bar{\pmb{Z}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\bar{\pmb{Z}}_G \label{eq:var_2p_reg_synth} \\
  \hat{\var}(\hat{Y}_{G,psynth,2p}) & := \hat{\bar{\pmb{Z}}}_G^t\hat{\pmb{\Sigma}}_{\hat{\pmb{\beta}}_{s_2}}\hat{\bar{\pmb{Z}}}_G
  + \hat{\pmb{\beta}}_{s_2}^t\hat{\Sigma}_{\hat{\bar{\pmb{Z}}}_G}\hat{\pmb{\beta}}_{s_2}  \label{eq:var_2p_reg_psynth}
\end{align}
\end{subequations}

%------------------------------------------------------------- %
% \textbf{Extended Synthetic and Extended Pseudo Synthetic Estimator}\par

The \textbf{\textit{extended}} \textbf{synthetic} and \textbf{\textit{extended}} \textbf{pseudo synthetic estimator}

% - explain extended estimator
% - mention boundary weight adjustment likewise global estimation

@@

%------------------------------------------------------------- %
\subsubsection{Application}

Small area estimations in the package \pkg{forestinventory} can be applied by \textit{additionally} specifying the optional argument \code{small_area}. The input dataset has therefore to be extended by an additional column that specifies for each each row (i.e. observation of a sample location) the respective small area in which it is located. The argument \code{small_area} requires a \code{list}-object that comprises 1) the name of the column specifiying the small area of each observation (\code{sa.col}), 2) a vector specifying the small area(s) for which estimations are desired (\code{areas}), and 3) the argument \code{unbiased} that controls which of the three available estimators should be applied. In addition to global estimation, the \code{summary} now also displays the sample sizes in the respective small area (\code{n1G} and \code{n2G}).

In order to apply the \textbf{pseudo small area estimator} (\textit{psmall}), we set \code{unbiased=TRUE} as well as the optional argument \code{psmall=TRUE}:
\begin{small}
<<>>=
psmall_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      boundary_weights = "boundary_weights",
                      small_area = list(sa.col = "smallarea", areas = c("A", "B"),
                                        unbiased = TRUE),
                      psmall = TRUE)
summary(psmall_2p)
@
\end{small}

The \textbf{extended pseudo small area estimator} (\textit{extpsynth}) can be applied by setting \code{unbiased=TRUE} and leaving the optional argument \code{psmall} to its default value of \code{FALSE}:
\begin{small}
<<>>=
extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                         phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                         boundary_weights = "boundary_weights",
                         small_area = list(sa.col = "smallarea", areas = c("A", "B"),
                                           unbiased = TRUE))
extpsynth_2p$estimation
@
\end{small}

Note that the point estimates and the design-based and external variances of both estimators are very similiar. However, the R$^2$s (\code{r.squared}) under the \textbf{extended pseudo small area estimator} are changing with the small area, while they are identical under the \textbf{pseudo small area estimator}. This is because under the \textbf{extended pseudo small area estimator}, the regression model is recalculated for each small area estimation after adding the indicator variable for the respective small area in the globally derived design matrices. In case of the \textbf{pseudo small area estimator}, the regression model stays the same for each small area estimation. Although the results of both estimators should always be close to each other, we recommend applying both estimators and compare the results afterwards in order to reveal unsuspected patterns in the data, particularly in the case of cluster sampling.\par

Setting the argument \code{unbiased=FALSE} applies the \textbf{pseudo synthetic estimator} to the selected small areas. Note that in the \code{grisons} dataset, all small areas possess at least 2 terrestrial observations ($n_{2,G} \geq 2$) and hence design-unbiased estimations are possible. Applying the pseudo synthetic estimator equals treating those small areas as if $n_{2,G}=0$ by ignoring the terrestrial information.
\begin{small}
<<>>=
psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      boundary_weights = "boundary_weights",
                      small_area = list(sa.col = "smallarea", areas = c("A", "B"),
                                        unbiased = FALSE))
psynth_2p$estimation
@
\end{small}

In this case, the variance of the pseudo synthetic estimator is almost only half as high as under the \textit{psmall} and \textit{extended psynth} estimator. The reason for this is that the pseudo synthetic estimator neglects the uncertainty of the regression model predictions in terms of incorporating the residual variation (compare equation  \ref{eq:var_2p_reg_psynth} to \ref{eq:var_2p_reg_psmall} and XXX). Consequently, relying on the synthetic estimation would in this case lead to a severe overestimation of the estimation precision.

>>>>

\section{Calculation of Confidence Intervals}
\label{sec:confint_calc}

@@
